Trade-offs among plant floral and reproductive traits determine interactions with floral visitors

### 2021-08-11 Draft/Manuscript.Rmd
edits on manuscript

header-includes:
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage{xcolor}
- \usepackage{float}
- \floatplacement{figure}{H}
- \usepackage[skip=5pt]{caption}
---

@@

\small

@@

\normalsize

Plant life strategies are often delimited by vegetative and physiological traits but little is known about how floral and reproductive traits drive these strategies, and in turn shape plant interactions with floral visitors. Here, we compiled 13 floral, 4 reproductive and 3 vegetative traits for 1,506 plant species from 28 plant-pollinator network studies across 18 different countries. We investigated the associations among these traits, pollinator visitation and the functional role of plant species within the networks (interaction frequency, normalized degree and specialization). We found that 51.8% of trait variation was explained by two independent axes that encompassed plant form and function. Specifically, the first axis indicated the presence of a trade-off between flower number and flower size (PC1, 26.72%). The second axis indicated a trade-off for the level of pollinator dependency (PC2, 25.08%). Although the main axes of trait variation did not fully explain pollinator visitation rates, different plant life strategies were associated with visitation rates and pollinator functional groups. Overall, the main traits that determined plant species’ functional roles were height, nectar concentration, pollen grains per flower, number of ovules, style length, selfing level and flower width. Our results highlight the need to consider plant reproductive and floral traits to improve understanding of plant life strategies and plant-pollinator interactions at broader spatial scales.

\vspace{5mm}
\bleft

@@

```{r, echo=FALSE, message=FALSE, cache=FALSE, results='hide', warning=FALSE, fig.height=10, fig.width=10,fig.pos = "!H",fig.cap= "Figure 1. Life history strategies for 1,236 plant species from 28 plant-pollinator networks studies across the first two axes of trait variation from a phylogenetically informed principal component analysis (pPCA). The solid arrows indicate the direction of the different quantitative traits (flower number, plant height, style length, flower size, ovule number and level of autonomous selfing) in the 2-d plane. The length of the arrows indicate the weight of the variables on each principal component and the labelled icons at their end represent the extreme form of the trait continuum. The dashed lines show the opposed direction of trait variation and the non-labelled icons at their end illustrate the opposing extreme of the continuum."}

@@

>>>>

header-includes:
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage{xcolor}
- \usepackage{float}
- \floatplacement{figure}{H}
- \usepackage[skip=3pt]{caption}
- \captionsetup[figure]{labelformat=empty}
- \usepackage{setspace}

---

@@

\small
\doublespacing

@@

\normalsize
\singlespacing

__Plant life strategies are often delimited by vegetative and physiological traits but little is known about how floral and reproductive traits drive these strategies, and in turn shape plant interactions with floral visitors. Here, we compiled 13 floral, 4 reproductive and 3 vegetative traits for 1,506 plant species from 28 plant-pollinator network studies across 18 different countries. We investigated the associations among these traits, pollinator visitation and the functional role of plant species within the networks (interaction frequency, normalized degree and specialization). We found that 51.8% of trait variation was explained by two independent axes that encompassed plant form and function. Specifically, the first axis indicated the presence of a trade-off between flower number and flower size (PC1, 26.72%). The second axis indicated a trade-off for the level of pollinator dependency (PC2, 25.08%). Although the main axes of trait variation did not fully explain pollinator visitation rates, different plant life strategies were associated with visitation rates and pollinator functional groups. Overall, the main traits that determined plant species’ functional roles were height, nectar concentration, pollen grains per flower, number of ovules, style length, selfing level and flower width. Our results highlight the need to consider plant reproductive and floral traits to improve understanding of plant life strategies and plant-pollinator interactions at broader spatial scales.__

\doublespacing
\vspace{5mm}
\bleft

@@

```{r, echo=FALSE, message=FALSE, cache=FALSE, results='hide', warning=FALSE, fig.height=10, fig.width=10,fig.pos = "t",fig.cap= "**Fig. 1 |** Life history strategies for 1,236 plant species from 28 plant-pollinator networks studies across the first two axes of trait variation from a phylogenetically informed principal component analysis (pPCA). The solid arrows indicate the direction of the different quantitative traits (flower number, plant height, style length, flower size, ovule number and level of autonomous selfing) in the 2-d plane. The length of the arrows indicate the weight of the variables on each principal component and the labelled icons at their end represent the extreme form of the trait continuum. The dashed lines show the opposed direction of trait variation and the non-labelled icons at their end illustrate the opposing extreme of the continuum."}

@@

\vspace*{-28mm}

```{r, echo=FALSE, message=FALSE, cache=FALSE, results='hide', warning=FALSE, fig.height=12, fig.width=18,fig.pos = "!H",fig.cap= "**Fig. 2 | Location of the different qualitative traits on the trait space.** The panel is composed by the traits that showed statistical association with the first two axes of trait variation: compatibility system (a), life form (b), lifespan (c), breeding system (d), flower shape (e) and flower symmetry (f)."}

#Load libraries
#LOAD LIBRARIES
#library(phytools)
#library(ape) #for phylogenetic distance
#library(dplyr) #data processing
#library(rtrees) #for phylogenetic distancelibrary(MASS)
#library(reshape2)
library(viridis) #COLOUR GGPLOT
#library(MASS)
library(ggplot2)
library(broman) #crayon colours
#library(cowplot)

# Theme for publication
theme_ms <- function(base_size=12, base_family="Helvetica") {
  (theme_bw(base_size = base_size, base_family = base_family)+
      theme(text=element_text(color="black"),
            axis.title=element_text( size = rel(1.75)),
            axis.text=element_text(size = rel(1.75), color = "black"),
            legend.title=element_text(face="bold"),
            legend.text=element_text(),
            legend.background=element_rect(fill="transparent"),
            legend.key.size = unit(0.55, 'lines'),
            panel.border=element_rect(color="black",size=1),
            panel.grid.minor.x =element_blank(),
            panel.grid.minor.y= element_blank(),
             panel.grid.major= element_blank(),
            plot.title = element_text(size = rel(1.65), face="bold")
      ))
}

#  plot_grid(Compatibility(PC),Life_form(PC),Life_span(PC),Breeding_syst(PC),Flower_sha(PC),Flowe#r_sym(PC),ncol=3,nrow=2,  align = "hv")

library(patchwork)
Compatibility(PC) + Life_form(PC) + Life_span(PC) + Breeding_syst(PC) + Flower_sha(PC) + Flower_sym(PC) + plot_layout(ncol=3)

\vspace{5mm}

### 2021-08-11 Draft/Manuscript.tex
edits on manuscript

% Add Supplementary Tables and Figures
% Code from https://stackoverflow.com/a/51337664
\newcommand{\beginsupplement}{
  \setcounter{table}{0}  
  \renewcommand{\thetable}{S\arabic{table}}
  \setcounter{figure}{0} 
  \renewcommand{\thefigure}{S\arabic{figure}}
}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\usepackage{float}
\floatplacement{figure}{H}
\usepackage[skip=5pt]{caption}
\usepackage{float}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

@@

\normalsize

Plant life strategies are often delimited by vegetative and
physiological traits but little is known about how floral and
reproductive traits drive these strategies, and in turn shape plant
interactions with floral visitors. Here, we compiled 13 floral, 4
reproductive and 3 vegetative traits for 1,506 plant species from 28
plant-pollinator network studies across 18 different countries. We
investigated the associations among these traits, pollinator visitation
and the functional role of plant species within the networks
(interaction frequency, normalized degree and specialization). We found
that 51.8\% of trait variation was explained by two independent axes
that encompassed plant form and function. Specifically, the first axis
indicated the presence of a trade-off between flower number and flower
size (PC1, 26.72\%). The second axis indicated a trade-off for the level
of pollinator dependency (PC2, 25.08\%). Although the main axes of trait
variation did not fully explain pollinator visitation rates, different
plant life strategies were associated with visitation rates and
pollinator functional groups. Overall, the main traits that determined
plant species' functional roles were height, nectar concentration,
pollen grains per flower, number of ovules, style length, selfing level
and flower width. Our results highlight the need to consider plant
reproductive and floral traits to improve understanding of plant life
strategies and plant-pollinator interactions at broader spatial scales.

\vspace{5mm} \bleft

@@

\begin{figure}
\centering
\includegraphics{output/figures/unnamed-chunk-1-1.pdf}
\caption{\label{fig:unnamed-chunk-1}Figure 1. Life history strategies for
1,236 plant species from 28 plant-pollinator networks studies across the
first two axes of trait variation from a phylogenetically informed
principal component analysis (pPCA). The solid arrows indicate the
direction of the different quantitative traits (flower number, plant
height, style length, flower size, ovule number and level of autonomous
selfing) in the 2-d plane. The length of the arrows indicate the weight
of the variables on each principal component and the labelled icons at
their end represent the extreme form of the trait continuum. The dashed
lines show the opposed direction of trait variation and the non-labelled
icons at their end illustrate the opposing extreme of the continuum.}
\end{figure}

@@

\begin{figure}[H]
\includegraphics{output/figures/unnamed-chunk-2-1} \caption{Location of the different qualitative traits that showed statistical association with the first two axes of trait variation. These different traits included: compatibility system (a), life form (b), lifespan (c), breeding system (d), flower shape (e) and flower symmetry (f).}\label{fig:unnamed-chunk-2}
\end{figure}

@@

\caption{\label{tab:unnamed-chunk-3}List of the 20 traits compiled in this study divided in quantitative and categorical traits. The different types of trait (vegetative ‘V’, floral ‘F’ and reproductive ‘R’), total records found for the 1,506 species and categories of the qualitative traits are also provided. }
\centering
\fontsize{10}{12}\selectfont
\begin{tabu} to \linewidth {>{}c>{\raggedright}X>{\centering}X>{}c>{\raggedright}X>{\raggedright}X>{\centering}X}
\toprule
\multicolumn{3}{c}{\textbf{Quantitative traits}} & \multicolumn{4}{c}{\textbf{Categorical traits}} \\
\cmidrule(l{3pt}r{3pt}){1-3} \cmidrule(l{3pt}r{3pt}){4-7}
\textbf{Type} & \textbf{Traits} & \textbf{Records} & \textbf{Type} & \textbf{Traits} & \textbf{Categories} & \textbf{Records}\\
\midrule
\textbf{V} & Plant height (m) & 1470 & \textbf{V} & Lifepan & \makecell[l]{Short-lived \\ Perennial} & 1466\\
\addlinespace
\textbf{F} & Flower width (mm) & 1472 & \textbf{V} & Life form & \makecell[l]{Herb \\ Shrub \\ Tree} & 1472\\
\addlinespace
\textbf{F} & Flower length (mm) & 1401 & \textbf{F} & Flower shape & \makecell[l]{Brush \\ Campanulate \\ Capitulum \\ Open \\ Papilionaceous \\ Tube} & 1458\\
\addlinespace
\textbf{F} & Inflorescence width (mm) & 1496 & \textbf{F} & Flower symmetry & \makecell[l]{Actinomorphic \\ Zygomorphic} & 1478\\
\addlinespace
\textbf{F} & Style length (mm) & 1497 & \textbf{F} & Nectar & \makecell[l]{Presence \\ Absence} & 1408\\
\addlinespace
\textbf{F} & Ovules per flower & 1457 & \textbf{R} & Autonomous selfing & \makecell[l]{None \\ Low \\ Medium \\ High} & 987\\
\addlinespace
\textbf{F} & Flowers per plant & 1468 & \textbf{R} & Compatibility system & \makecell[l]{Self-incomp. \\ Partially self-comp. \\ Self-comp.} & 1253\\
\addlinespace
\textbf{F} & Nectar ($\mu$l) & 531 & \textbf{R} & Breeding system & \makecell[l]{Hermaphrodite \\ Monoecious \\ Dioecious} & 1489\\
\addlinespace
\textbf{F} & Nectar (mg) & 455 & \textbf{} &  &  & \\
\addlinespace
\textbf{F} & Nectar concentration ($\%$) & 558 & \textbf{} &  &  & \\
\addlinespace
\textbf{F} & Pollen grains per flower & 533 & \textbf{} &  &  & \\
\addlinespace
\textbf{R} & Autonomous selfing (fruit set) & 992 & \textbf{} &  &  & \\
\bottomrule
\end{tabu}
\end{table}

@@

\begin{figure}
\centering
\includegraphics{output/figures/unnamed-chunk-4-1.pdf}
\caption{\label{fig:unnamed-chunk-4}Life history strategies for 1,236 plant
species from 29 plant-pollinator networks studies across the first two
axes of variation from a phylogenetically informed principal component
analysis (pPCA). The solid arrows indicate the direction and weights of
the six quantitative traits (flower number, plant height, style length,
flower size, ovule number and autonomous selfing level) and the labelled
icons at their end represent the extreme form of the trait continuum.
The dashed lines indicate the opposed direction of trait variation and
the non-labelled icons at their end illustrate the other extreme of the
continuum.}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics{output/figures/unnamed-chunk-5-1.pdf}
\caption{\label{fig:unnamed-chunk-5}Location of the qualitative traits in
the trait space for traits that showed a statistical association with
the main axis of trait variation. These different traits included:
compatibility system (a), life form (b), lifespan (c), breeding system
(d), flower shape (e) and flower symmetry (f).}
\end{figure}

@@

\includegraphics{output/figures/unnamed-chunk-6-1.pdf}

>>>>

% Add Supplementary Tables and Figures
% Code from https://stackoverflow.com/a/51337664
\newcommand{\beginsupplement}{
  \setcounter{table}{0}  
  \renewcommand{\thetable}{S\arabic{table}}
  \setcounter{figure}{0} 
  \renewcommand{\thefigure}{S\arabic{figure}}
}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\usepackage{float}
\floatplacement{figure}{H}
\usepackage[skip=3pt]{caption}
\captionsetup[figure]{labelformat=empty}
\usepackage{setspace}
\usepackage{float}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

@@

\normalsize
\singlespacing

\textbf{Plant life strategies are often delimited by vegetative and
physiological traits but little is known about how floral and
reproductive traits drive these strategies, and in turn shape plant
interactions with floral visitors. Here, we compiled 13 floral, 4
reproductive and 3 vegetative traits for 1,506 plant species from 28
plant-pollinator network studies across 18 different countries. We
investigated the associations among these traits, pollinator visitation
and the functional role of plant species within the networks
(interaction frequency, normalized degree and specialization). We found
that 51.8\% of trait variation was explained by two independent axes
that encompassed plant form and function. Specifically, the first axis
indicated the presence of a trade-off between flower number and flower
size (PC1, 26.72\%). The second axis indicated a trade-off for the level
of pollinator dependency (PC2, 25.08\%). Although the main axes of trait
variation did not fully explain pollinator visitation rates, different
plant life strategies were associated with visitation rates and
pollinator functional groups. Overall, the main traits that determined
plant species' functional roles were height, nectar concentration,
pollen grains per flower, number of ovules, style length, selfing level
and flower width. Our results highlight the need to consider plant
reproductive and floral traits to improve understanding of plant life
strategies and plant-pollinator interactions at broader spatial scales.}

\doublespacing
\vspace{5mm} \bleft

@@

\begin{figure}
\centering
\includegraphics{output/figures/unnamed-chunk-1-1.pdf}
\caption{\label{fig:unnamed-chunk-1}\textbf{Fig. 1 \textbar{}} Life history
strategies for 1,236 plant species from 28 plant-pollinator networks
studies across the first two axes of trait variation from a
phylogenetically informed principal component analysis (pPCA). The solid
arrows indicate the direction of the different quantitative traits
(flower number, plant height, style length, flower size, ovule number
and level of autonomous selfing) in the 2-d plane. The length of the
arrows indicate the weight of the variables on each principal component
and the labelled icons at their end represent the extreme form of the
trait continuum. The dashed lines show the opposed direction of trait
variation and the non-labelled icons at their end illustrate the
opposing extreme of the continuum.}
\end{figure}

@@

\vspace*{-28mm}

\begin{figure}
\centering
\includegraphics{output/figures/unnamed-chunk-2-1.pdf}
\caption{\label{fig:unnamed-chunk-2}\textbf{Fig. 2 \textbar{} Location of
the different qualitative traits on the trait space.} The panel is
composed by the traits that showed statistical association with the
first two axes of trait variation: compatibility system (a), life form
(b), lifespan (c), breeding system (d), flower shape (e) and flower
symmetry (f).}
\end{figure}

\vspace{5mm}

\begin{figure}[H]
\includegraphics{output/figures/unnamed-chunk-3-1} \caption{Location of the different qualitative traits that showed statistical association with the first two axes of trait variation. These different traits included: compatibility system (a), life form (b), lifespan (c), breeding system (d), flower shape (e) and flower symmetry (f).}\label{fig:unnamed-chunk-3}
\end{figure}

@@

\caption{\label{tab:unnamed-chunk-4}List of the 20 traits compiled in this study divided in quantitative and categorical traits. The different types of trait (vegetative ‘V’, floral ‘F’ and reproductive ‘R’), total records found for the 1,506 species and categories of the qualitative traits are also provided. }
\centering
\fontsize{10}{12}\selectfont
\begin{tabu} to \linewidth {>{}c>{\raggedright}X>{\centering}X>{}c>{\raggedright}X>{\raggedright}X>{\centering}X}
\toprule
\multicolumn{3}{c}{\textbf{Quantitative traits}} & \multicolumn{4}{c}{\textbf{Categorical traits}} \\
\cmidrule(l{3pt}r{3pt}){1-3} \cmidrule(l{3pt}r{3pt}){4-7}
\textbf{Type} & \textbf{Traits} & \textbf{Records} & \textbf{Type} & \textbf{Traits} & \textbf{Categories} & \textbf{Records}\\
\midrule
\textbf{V} & Plant height (m) & 1470 & \textbf{V} & Lifepan & \makecell[l]{Short-lived \\ Perennial} & 1466\\
\addlinespace
\textbf{F} & Flower width (mm) & 1472 & \textbf{V} & Life form & \makecell[l]{Herb \\ Shrub \\ Tree} & 1472\\
\addlinespace
\textbf{F} & Flower length (mm) & 1401 & \textbf{F} & Flower shape & \makecell[l]{Brush \\ Campanulate \\ Capitulum \\ Open \\ Papilionaceous \\ Tube} & 1458\\
\addlinespace
\textbf{F} & Inflorescence width (mm) & 1496 & \textbf{F} & Flower symmetry & \makecell[l]{Actinomorphic \\ Zygomorphic} & 1478\\
\addlinespace
\textbf{F} & Style length (mm) & 1497 & \textbf{F} & Nectar & \makecell[l]{Presence \\ Absence} & 1408\\
\addlinespace
\textbf{F} & Ovules per flower & 1457 & \textbf{R} & Autonomous selfing & \makecell[l]{None \\ Low \\ Medium \\ High} & 987\\
\addlinespace
\textbf{F} & Flowers per plant & 1468 & \textbf{R} & Compatibility system & \makecell[l]{Self-incomp. \\ Partially self-comp. \\ Self-comp.} & 1253\\
\addlinespace
\textbf{F} & Nectar ($\mu$l) & 531 & \textbf{R} & Breeding system & \makecell[l]{Hermaphrodite \\ Monoecious \\ Dioecious} & 1489\\
\addlinespace
\textbf{F} & Nectar (mg) & 455 & \textbf{} &  &  & \\
\addlinespace
\textbf{F} & Nectar concentration ($\%$) & 558 & \textbf{} &  &  & \\
\addlinespace
\textbf{F} & Pollen grains per flower & 533 & \textbf{} &  &  & \\
\addlinespace
\textbf{R} & Autonomous selfing (fruit set) & 992 & \textbf{} &  &  & \\
\bottomrule
\end{tabu}
\end{table}

@@

\begin{figure}
\centering
\includegraphics{output/figures/unnamed-chunk-5-1.pdf}
\caption{\label{fig:unnamed-chunk-5}Life history strategies for 1,236 plant
species from 29 plant-pollinator networks studies across the first two
axes of variation from a phylogenetically informed principal component
analysis (pPCA). The solid arrows indicate the direction and weights of
the six quantitative traits (flower number, plant height, style length,
flower size, ovule number and autonomous selfing level) and the labelled
icons at their end represent the extreme form of the trait continuum.
The dashed lines indicate the opposed direction of trait variation and
the non-labelled icons at their end illustrate the other extreme of the
continuum.}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics{output/figures/unnamed-chunk-6-1.pdf}
\caption{\label{fig:unnamed-chunk-6}Location of the qualitative traits in
the trait space for traits that showed a statistical association with
the main axis of trait variation. These different traits included:
compatibility system (a), life form (b), lifespan (c), breeding system
(d), flower shape (e) and flower symmetry (f).}
\end{figure}

@@

\includegraphics{output/figures/unnamed-chunk-7-1.pdf}
